---
title: "webR with Quarto HTML Standalone Document Proof of Concept"
subtitle: "Experiments with an Interactive Quarto Document using webR v0.1.0"
author: "JJB leaning heavily on webR authors"
engine: knitr
execute:
  echo: true
  error: true
embed-resources: true
format:
    html: 
      toc: true
editor: source
---

# Demo

## Background

The purpose of this document is to explore how WebR can be embedded in a 
Quarto Document for the purposes of teaching _R_. 

- WebR Website: <https://docs.r-wasm.org/webr/latest/>
- WebR GitHub: <https://github.com/r-wasm/webr/>

## Setup 

See the <https://github.com/coatless-r-n-d/webR-quarto-demos> for source.

```{=html}
<link rel="stylesheet" href="https://cdnjs.cloudflare.com/ajax/libs/codemirror/6.65.7/codemirror.min.css">
<style>
  .CodeMirror pre {
    background-color: unset !important;
  }
  .btn-webr {
    background-color: #EEEEEE;
    border-bottom-left-radius: 0;
    border-bottom-right-radius: 0;
  }
</style>
<script src="https://cdnjs.cloudflare.com/ajax/libs/codemirror/6.65.7/codemirror.min.js"></script>
<script src="https://cdnjs.cloudflare.com/ajax/libs/codemirror/6.65.7/mode/r/r.js"></script>
<script type="module">
  import { WebR } from "https://webr.r-wasm.org/v0.1.0/webr.mjs";
  globalThis.webR = new WebR({
    SW_URL: "/"
  });
  await globalThis.webR.init();
  globalThis.webRCodeShelter = await new globalThis.webR.Shelter();
  document.querySelectorAll(".btn-webr").forEach((btn) => {
    btn.innerText = "Run code";
    btn.disabled = false;
  });
</script>
```

```{r}
#| results: asis
#| echo: false
webr_counter = 0

cat("importScripts('https://webr.r-wasm.org/v0.1.0/webr-worker.js');", file = "webr-worker.js")
cat("importScripts('https://webr.r-wasm.org/v0.1.0/webr-serviceworker.js');", file = "webr-serviceworker.js")

webr_editor = function(code = I(encodeString(code, quote = '`')), width, height) { 
webr_counter <<- webr_counter + 1

output = glue::glue('
<button class="btn btn-default btn-webr" disabled type="button"  id="webr-run-button-{{ webr_counter }}">Loading webR...</button>
<div id="webr-editor-{{ webr_counter }}"></div>
<div id="webr-code-output-{{ webr_counter }}"><pre style="visibility: hidden"></pre></div>
<script type="module">
  const runButton = document.getElementById("webr-run-button-{{ webr_counter }}");
  const outputDiv = document.getElementById("webr-code-output-{{ webr_counter }}");
  const editorDiv = document.getElementById("webr-editor-{{ webr_counter }}");

  const editor = CodeMirror((elt) => {
    elt.style.border = "1px solid #eee";
    elt.style.height = "auto";
    editorDiv.append(elt);
  },{
    value: {{code}},
    lineNumbers: true,
    mode: "r",
    theme: "light default",
    viewportMargin: Infinity,
  });

  runButton.onclick = async () => {
    runButton.disabled = true;
    let canvas = undefined;
    await globalThis.webR.init();
    await webR.evalRVoid("canvas(width={{width}}, height={{height}})");
    const result = await webRCodeShelter.captureR(editor.getValue(), {
      withAutoprint: true,
      captureStreams: true,
      captureConditions: false,
      env: webR.objs.emptyEnv,
    });
    try {
      await webR.evalRVoid("dev.off()");
      const out = result.output.filter(
        evt => evt.type == "stdout" || evt.type == "stderr"
      ).map((evt) => evt.data).join("\\n");

      const msgs = await webR.flush();
      msgs.forEach(msg => {
        if (msg.type === "canvasExec"){
          if (!canvas) {
            canvas = document.createElement("canvas");
            canvas.setAttribute("width", 2 * {{width}});
            canvas.setAttribute("height", 2 * {{height}});
            canvas.style.width="700px";
            canvas.style.display="block";
            canvas.style.margin="auto";
          }
          Function(`this.getContext("2d").${msg.data}`).bind(canvas)();
        }
      });

      outputDiv.innerHTML = "";
      const pre = document.createElement("pre");
      if (/\\S/.test(out)) {
        const code = document.createElement("code");
        code.innerText = out;
        pre.appendChild(code);
      } else {
        pre.style.visibility = "hidden";
      }
      outputDiv.appendChild(pre);

      if (canvas) {
        const p = document.createElement("p");
        p.appendChild(canvas);
        outputDiv.appendChild(p);
      }
    } finally {
      webRCodeShelter.purge();
      runButton.disabled = false;
    }
  }

  await globalThis.webR.init();
  runButton.innerText = "Run code";
  runButton.disabled = false;
</script>
', .open = "{{", .close = "}}")
}
```

```{r}
#| echo: false
knitr::knit_engines$set(webr = function(options) {
  code = paste(options$code, collapse = "\n")
  w = knitr::opts_current$get('fig.width') * 72
  h = knitr::opts_current$get('fig.height') * 72
  options$results = 'asis'
  
  form = webr_editor(code = I(encodeString(code, quote = '`')), width = w, height = h)
  
  form
}
)
```

## Exploration

Next, let's look at a few features of the language

### Linear Regression

We'll first start with the WebR team's demo example or the statistician way of
saying, "Hello world!"... Aka linear regression:

```{webr}
fit = lm(mpg ~ am, data=mtcars)
summary(fit)
```

### Retrieving prior objects

Each WebR cell appears to be connected to each other. Thus, we can access the
`fit` outcome:

```{webr}
coef(fit)
```

```{webr}
anova(fit)
```

### Mixing active and non-active _R_ code

For _if-else_ statements, we have:

```r
if (...) {
  # Statements for TRUE
} else {
  # Statements for FALSE
}
```

- `...` denotes a condition (either `TRUE` or `FALSE`)
- If `TRUE`, then run the statements inside `{}`
- Else, `FALSE`, carry on with your day. 

How could we modify `temperature` to have the `if` statement print `"Hot!"`? 

```{webr}
# Let's classify 
temperature = 60

if (temperature > 76) {
    print("Hot!")
} else {
    print("Cold!")
}
```


### Summarize Data

Glancing at data frames yields: 

```{webr}
summary(mtcars)
```

### Errors and Warnings


```{webr}
stop("What happens if an error is present?")
```

```{webr}
warning("You shouldn't be here...")
```


### Base graphics

Graphing with base R 

```{webr}
plot(pressure)
```

More advanced base R graphing... 

```{webr}
x1 = seq(0, 1, length = 20)
y1 = rep(0, 20)
x2 = rep(0, 20)
y2 = seq(0.75, 0, length = 20)

plot(0, type = "n",
     axes = FALSE, ylab = "", xlab = "", 
     xlim = c(0, 1), ylim = c(0, 0.75), asp = 1,
     main = "Straight Lines as a Curve")

segments(x1, y1, x2, y2)
box(col = "grey")
```

### ggplot2 Graphics

Next, we look at using `ggplot2` graphics. By default, the `ggplot2` package
is not available as it is _dependency_ heavy. 

```{=html}
<details>
<summary>
Package installation for `ggplot2` given by `webr::install("ggplot2")`
</summary>
Downloading webR package: cli
Downloading webR package: glue
Downloading webR package: gtable
Downloading webR package: isoband
Downloading webR package: rlang
Downloading webR package: lifecycle
Downloading webR package: MASS
Downloading webR package: lattice
Downloading webR package: nlme
Downloading webR package: Matrix
Downloading webR package: mgcv
Downloading webR package: farver
Downloading webR package: labeling
Downloading webR package: colorspace
Downloading webR package: munsell
Downloading webR package: R6
Downloading webR package: RColorBrewer
Downloading webR package: viridisLite
Downloading webR package: scales
Downloading webR package: fansi
Downloading webR package: magrittr
Downloading webR package: utf8
Downloading webR package: vctrs
Downloading webR package: pillar
Downloading webR package: pkgconfig
Downloading webR package: tibble
Downloading webR package: withr
Downloading webR package: ggplot2
</details>
```

```{webr}
# Install non-base R packages
webr::install("ggplot2")
# Load non-base packages like normal
library("ggplot2")
p = ggplot(mpg, aes(class, hwy))
p + geom_boxplot()
```



